Network Trait marker for working memory deficit in Schizophrenia: a multi site study

### 2016-03-06 manuscript.tex
rest
rest

There has been abundant imaging studies on WM deficits in Schizophrenia, but the differences among the tasks and material make it difficult to make a conclusion.  Some studies have found evidence for different patterns of functional connectivity during different working memory task conditions (e.g., as a function of load, stimulus type, or task phase). Such findings suggest that functional connectivity changes could reflect differences in task engagement or responsivity of brain networks to modulation,rather than stable changes that persist across all task states. On the other hand, working memory itself is a complex system which consist of different components. It is widely agreed that WM involves several different component processes. One popular and original model of WM is a system encoding and maintaining modality-specific short-term memory and central executive component to manipulate the information. Some studies have found both impairments in the modality-specific perception component and the later manipulation component. There have been evidence of multi-modality deficits: visual spatial, visual object, verbal, and other types of working memory. In a recent meta analysis, the author draw attention to a consistent and restricted "core network" emerged from conjunctions across analyses of specific task designs and contrasts. This distributed network was believed to be active in WM task ignoring the task type , stimulus type and WM load and may be act as a base part in WM. We restrict our analysis in these core networks.

@@

Two types of 3T MRI scanners (four Siemens, three General Electric) were used at the participating centers (details in table \#). To ensure equivalent data acquisition protocol and high quality imaging data, the scanning parameters of the functional scans at each of the six centers were set up by an experienced researcher before data acquisition. An echo planar imaging sequence was used to obtain the functional images, the parameters were as follows: 30 axial slices, TR = 2000 ms, TE = 30 ms, matrix = 64 × 64, flip angle = 90°, FOV = 240*240 mm 2, slice thickness = 4 mm, gap = 0.4 mm. A total of 250 brain volumes were collected, resulting in a total scan time of 500s. The MRI scan sequences and parameters for each center are listed in eTable 1. To be noticed, the time point number of the images from ZMD is 180 and is different from the 240 of the other sites. We did not exclude this site since we are eager to keep as much data and sites as possible, to support the validation of the multi-site analysis. The multi-site statisical stratergy we used is robust to various of volum numbers in different sites.

@@

The iamges were preprocessed with a based in-house software: Brainnetome Toolkit which utilized Statistical Parametric Mapping SPM8 (\url{http://www.fil.ion.ucl.ac.uk/spm}). The pipeline includes the following steps: The first ten images were deleted for the signal equilibration. The remaining iamges were conducted for slice acquisition correction and head motion correction.  The fMRI data which had
less than 3.0mm of head motion and 3.0° of angular rotation were included.  Moreover, the mean frame-wise displacement (FD) was computed by averaging FD\_i from every time point for each subject. There were no differences for the mean FD between groups ( t = 0.413, p = 0.682) (Table 1). Then the fMRI images were normalized to the standard Montreal Neurological Institute (MNI) template provided by SPM and resample to the 3-mm isotropic voxels. Artefacts due to changes in global, ventricle and white matter signals, residual motion were removed using voxel-wise regression. A temporal filter (0.01 Hz \textless f \textless 0.08 Hz) were used to reduce the low-frequency drift and physiological high frequency respiratory and cardiac noise. Finally, the data was smoothed with an isotropic Gaussian kernel of 6 mm full-width at half-maximum.

@@

The ROIs in this study were extracted from a previous meta analysis by Simon, in which they find regions that commonly acvite during various working memory tasks. A consistent and restricted "core" network emerged from conjunctions across analyses of specific task designs and contrasts: task effects for n-back and Sternberg tasks, verbal and non-verbal tasks, load effects and all three task components (encoding, maintenance, recall). These regions includs dosal area 44 , anterior insula, (pre-) SMA, and IPS, bilaterally. The peak coordinates are listed in table \#. A 5mm radis sphrere ROIs are created form these peak coordinates in MNI standard space.

@@

For every single site separately, the diagnose effect on each FC value was assessed using a two-sample t-test comparing images between SZ and NC from each site. A second level random effects meta analysis was used to pooling the single site statistics, in order to assess the replication of the effects from the entire dataset. The effect sizes were calculated using Hedge’s g, which provides an unbiased standardized mean difference that incorporates a correction for small sample sizes. Hedge’s g was calculated using each center’s t-statistic and the corresponding sample size. All significance thresholds were set to p \textless 0.001 FDR-corrected for multiple comparison. The meta-analysis was conducted in R with meta-for packages.

@@

Working memory is the result of various combinations of processes, no processes (and correspondingly no brain structures)are unique or specific to working memory( Eriksson, 2015). Many brain regions interact during working memory and include "executive" regions in the PFC, parietal cortex, and basal ganglia, as well as regions specialized for processing the particular representations to be maintained, such as the fusiform face area for maintaining face information. Persistent neural activity in various brain regions accompanies working memory and is functionally necessary for maintenance and integration of information in working memory.

The organization of human WM has long been the topic of psychological models (Atkinson and Shiffrin, 1968; Hebb, 1949), with maybe the most influential having been proposed by Baddeley and Hitch (1974). In Rottschy,2012, the WM network was identified by functional neuroimaging using quantiattive coordiante-based meta-analysis over almost 200 individual experiments. By pooling various working memory tasks, a main network was identified which mainly comprise the fronto-parietal network. By eliminating the effect of specific task designs and contrasts, a more restricted "core" network emerged from conjunctions analyses. A core network independently of the specific aspects and task features was identified using conjunction analysis. This network mainly comprise the dorsal area 44, anterior insula, (pre-)SMA and IPS. The dorsal areas 44/45 and the pre-SMA are part of the phonological loop, a subsystem that response for verbal working memory material maintenance. There has been evidence that the dorsal region of Broca's area is active only during the first part of the delay period, and is involved in the formation of an articulatory rehearsal program. Generally, the same brain regions dedicated to sensory processing are believed to store sensory information during delay periods and working-memory task performance. A conjunction of verbal vs. non-verbal material in Rottschy's study revealed that the BA44/45 area may also involed in non-verbal WM tasks, that it may not be a modality-specific area in working memory. The reduced functional connectivity may underpin the verbal working memory deficits in schizophrenia. The previous research has established that the PFC is causally involved in normal working  memory functioning. However, there is yet no consensus on the details of the functional organization of the PFC. The absent of DLPFC, which is believed to an neuro agent for the central executve module in Baddeley and Hitch's model in Rottschy's core network, may be due to some working memory task requires little manipulation process of the memory content. In fact, the lateral PFC clusters in the main network was subdivided to two, that the abstraction level of goals and task rules are suggested to peak in rostral PFC and decrease to the caudal part. The caudal LPFC register a working memory load effect while the rostral part was not. A meta analysis has find consistent activation in bilateral mid-ventrolateral prefrontal coretex or frontal operculum(BA45,47) whin N-back studies. (Owen, 2005) In this study, three subsets of the N-back tasks revealed similar activation pattern implicating prefrontal, premotor, and posterior parietal cortex, which suggested a core modality independent working memory network. The inhibitory processes apear to be mediated by area 45 (left lateral prefrontal structures) (Jonides,1998) in working memory tasks.  For  example,  Owen,Petrides and their colleagues [6,7] proposed that the mid-ventrolateral region (Brodmann’s area [BA] 45/47) supports the organization of response sequences based on information  retrieved  from  posterior  areas,  whereas  the mid-dorsolateral  region  (BA 9/46)  supports  the  active manipulation or monitoring of information within working memory(Patricia A Carpenter, 2000).

@@

>>>>

There has been abundant imaging studies on WM deficits in Schizophrenia, but the differences among the tasks and material make it difficult to make a conclusion.  Some studies have found evidence for different patterns of functional connectivity during different working memory task conditions (e.g., as a function of load, stimulus type, or task phase). Such findings suggest that functional connectivity changes could reflect differences in task engagement or responsivity of brain networks to modulation,rather than stable changes that persist across all task states. On the other hand, working memory itself is a complex system which consist of different components. It is widely agreed that WM involves several different component processes. One popular and original model of WM is a system encoding and maintaining modality-specific short-term memory and central executive component to manipulate the information. Some studies have found both impairments in the modality-specific perception component and the later manipulation component. There have been evidence of multi-modality deficits: visual spatial, visual object, verbal, and other types of working memory. In a recent meta analysis, the author draw attention to a consistent and restricted ``core network'' emerged from conjunctions across analyses of specific task designs and contrasts. This distributed network was believed to be active in WM task ignoring the task type , stimulus type and WM load and may be act as a base part in WM. We restrict our analysis in these core networks.

@@

Two types of 3T MRI scanners (four Siemens, three General Electric) were used at the participating centers (details in table \#). To ensure equivalent data acquisition protocol and high quality imaging data, the scanning parameters of the functional scans at each of the six centers were set up by an experienced researcher before data acquisition. An echo planar imaging sequence was used to obtain the functional images, the parameters were as follows: 30 axial slices, TR = 2000 ms, TE = 30 ms, matrix = 64 × 64, flip angle = $90^{\circ}$, FOV = $240*240 mm^{2}$, slice thickness = 4 mm, gap = 0.4 mm. A total of 250 brain volumes were collected, resulting in a total scan time of 500s. The MRI scan sequences and parameters for each center are listed in eTable 1. To be noticed, the time point number of the images from ZMD is 180 and is different from the 240 of the other sites. We did not exclude this site since we are eager to keep as much data and sites as possible, to support the validation of the multi-site analysis. The multi-site statisical stratergy we used is robust to various of volum numbers in different sites.

@@

The iamges were preprocessed with a based in-house software: Brainnetome Toolkit which utilized Statistical Parametric Mapping SPM8 (\url{http://www.fil.ion.ucl.ac.uk/spm}). The pipeline includes the following steps: The first ten images were deleted for the signal equilibration. The remaining iamges were conducted for slice acquisition correction and head motion correction.  The fMRI data which had
less than 3.0mm of head motion and $3.0^{\circ}$ of angular rotation were included.  Moreover, the mean frame-wise displacement (FD) was computed by averaging \(FD_{i}\) from every time point for each subject. There were no differences for the mean FD between groups ( t = 0.413, p = 0.682) (Table 1). Then the fMRI images were normalized to the standard Montreal Neurological Institute (MNI) template provided by SPM and resample to the 3-mm isotropic voxels. Artefacts due to changes in global, ventricle and white matter signals, residual motion were removed using voxel-wise regression. A temporal filter (0.01 Hz \textless f \textless 0.08 Hz) were used to reduce the low-frequency drift and physiological high frequency respiratory and cardiac noise. Finally, the data was smoothed with an isotropic Gaussian kernel of 6 mm full-width at half-maximum.

@@

The ROIs in this study were extracted from a previous meta analysis by Simon, in which they find regions that commonly acvite during various working memory tasks. A consistent and restricted ``core'' network emerged from conjunctions across analyses of specific task designs and contrasts: task effects for n-back and Sternberg tasks, verbal and non-verbal tasks, load effects and all three task components (encoding, maintenance, recall). These regions includs dosal area 44 , anterior insula, (pre-) SMA, and IPS, bilaterally. The peak coordinates are listed in table \#. A 5mm radis sphrere ROIs are created form these peak coordinates in MNI standard space.

@@

For every single site separately, the diagnose effect on each FC value was assessed using a two-sample t-test comparing images between SZ and NC from each site. A second level random effects meta analysis was used to pooling the single site statistics, in order to assess the replication of the effects from the entire dataset. The effect sizes were calculated using Hedge's g, which provides an unbiased standardized mean difference that incorporates a correction for small sample sizes. Hedge's g was calculated using each center's t-statistic and the corresponding sample size. All significance thresholds were set to p \textless 0.001 FDR-corrected for multiple comparison. The meta-analysis was conducted in R with meta-for packages.

@@

Working memory is the result of various combinations of processes, no processes (and correspondingly no brain structures)are unique or specific to working memory( Eriksson, 2015). Many brain regions interact during working memory and include ``executive'' regions in the PFC, parietal cortex, and basal ganglia, as well as regions specialized for processing the particular representations to be maintained, such as the fusiform face area for maintaining face information. Persistent neural activity in various brain regions accompanies working memory and is functionally necessary for maintenance and integration of information in working memory.

The organization of human WM has long been the topic of psychological models (Atkinson and Shiffrin, 1968; Hebb, 1949), with maybe the most influential having been proposed by Baddeley and Hitch (1974). In Rottschy,2012, the WM network was identified by functional neuroimaging using quantiattive coordiante-based meta-analysis over almost 200 individual experiments. By pooling various working memory tasks, a main network was identified which mainly comprise the fronto-parietal network. By eliminating the effect of specific task designs and contrasts, a more restricted ``core" network emerged from conjunctions analyses. A core network independently of the specific aspects and task features was identified using conjunction analysis. This network mainly comprise the dorsal area 44, anterior insula, (pre-)SMA and IPS. The dorsal areas 44/45 and the pre-SMA are part of the phonological loop, a subsystem that response for verbal working memory material maintenance. There has been evidence that the dorsal region of Broca's area is active only during the first part of the delay period, and is involved in the formation of an articulatory rehearsal program. Generally, the same brain regions dedicated to sensory processing are believed to store sensory information during delay periods and working-memory task performance. A conjunction of verbal vs. non-verbal material in Rottschy's study revealed that the BA44/45 area may also involed in non-verbal WM tasks, that it may not be a modality-specific area in working memory. The reduced functional connectivity may underpin the verbal working memory deficits in schizophrenia. The previous research has established that the PFC is causally involved in normal working  memory functioning. However, there is yet no consensus on the details of the functional organization of the PFC. The absent of DLPFC, which is believed to an neuro agent for the central executve module in Baddeley and Hitch's model in Rottschy's core network, may be due to some working memory task requires little manipulation process of the memory content. In fact, the lateral PFC clusters in the main network was subdivided to two, that the abstraction level of goals and task rules are suggested to peak in rostral PFC and decrease to the caudal part. The caudal LPFC register a working memory load effect while the rostral part was not. A meta analysis has find consistent activation in bilateral mid-ventrolateral prefrontal coretex or frontal operculum(BA45,47) whin N-back studies. (Owen, 2005) In this study, three subsets of the N-back tasks revealed similar activation pattern implicating prefrontal, premotor, and posterior parietal cortex, which suggested a core modality independent working memory network. The inhibitory processes apear to be mediated by area 45 (left lateral prefrontal structures) (Jonides,1998) in working memory tasks.  For  example,  Owen,Petrides and their colleagues [6,7] proposed that the mid-ventrolateral region (Brodmann’s area [BA] 45/47) supports the organization of response sequences based on information  retrieved  from  posterior  areas,  whereas  the mid-dorsolateral  region  (BA 9/46)  supports  the  active manipulation or monitoring of information within working memory(Patricia A Carpenter, 2000).

@@

\subsection*{Limitations}
The present findings must be considered with respect to serveral limitations.
